---
title: "Report Code"
author: "Sebastian Kirkpatrick"
date: "2024-02-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("tidyverse")
library("viridis")
```

```{r}
szn <- read_csv("fangraphs_season_level.csv")

pitches <- read_csv("savant_pitch_level.csv")
```

# Role Clustering

```{r}
szn_filt <- szn |>
  filter(IP > 25 & G > 15) |>
  select_if(~!is.logical(.)) |>
  mutate(P_per_G = Pitches/G,
         Full_IP = trunc(IP),
         Partial_Outs = IP - trunc(IP),
         Outs = 3 * Full_IP + 10 * Partial_Outs,
         O_per_G = Outs/G, 
         Innings = Outs/3)
```

```{r}
tree <- szn_filt |>
  select(O_per_G, gmLI) |>
  scale(center = T, scale = T) |>
  dist(method = "euclidean") |>
  hclust(method="complete")
```

```{r}
plot(tree)
```

## Cluster Summaries

```{r}
szn_filt |>
  select(NameASCII, SV, P_per_G, O_per_G, Innings, gmLI) |>
  mutate(cluster = cutree(tree,k=7)) |>
  group_by(cluster) |>
  summarize(mean_P = mean(P_per_G), 
            mean_O = mean(O_per_G), 
            mean_I = mean(Innings), 
            mean_gmLI = mean(gmLI), 
            mean_SV = mean(SV), 
            count = n())
```

## Full Clusters

```{r}
szn_filt |>
  select(NameASCII, Season, Role, SV, P_per_G, O_per_G, Innings, gmLI) |>
  mutate(cluster = cutree(tree,k=7)) |>
  filter(cluster == 1)
```

```{r}
szn_filt |>
  select(NameASCII, Season, Role, SV, P_per_G, O_per_G, Innings, gmLI) |>
  mutate(cluster = cutree(tree,k=7)) |>
  filter(cluster == 2)
```

```{r}
szn_filt |>
  select(NameASCII, Season, Role, SV, P_per_G, O_per_G, Innings, gmLI) |>
  mutate(cluster = cutree(tree,k=7)) |>
  filter(cluster == 3)
```

```{r}
szn_filt |>
  select(NameASCII, Season, Role, SV, P_per_G, O_per_G, Innings, gmLI) |>
  mutate(cluster = cutree(tree,k=7)) |>
  filter(cluster == 4)
```

```{r}
szn_filt |>
  select(NameASCII, Season, Role, SV, P_per_G, O_per_G, Innings, gmLI) |>
  mutate(cluster = cutree(tree,k=7)) |>
  filter(cluster == 5)
```

```{r}
szn_filt |>
  select(NameASCII, Season, Role, SV, P_per_G, O_per_G, Innings, gmLI) |>
  mutate(cluster = cutree(tree,k=7)) |>
  filter(cluster == 6)
```

```{r}
szn_filt |>
  select(NameASCII, Season, Role, SV, P_per_G, O_per_G, Innings, gmLI) |>
  mutate(cluster = cutree(tree,k=7)) |>
  filter(cluster == 7)
```

# Arsenal Percent & Stuff + Correlations

## Starters

### Pitch Percentages

```{r}
szn_pitch <- szn |>
  replace(is.na(szn),0) |>
  filter(Role == "SP", G > 15) |>
  select(4, 6, 9, 12, 97, 170:172, 174:177, 179, 180) |>
  rename(
    "Fastball" = 6, 
    "Cutter" = 7, 
    "Split_Finger" = 8, 
    "Sinker" = 9, 
    "Slider" = 10, 
    "Curveball" = 11, 
    "Knuckle_Curve" = 12, 
    "Changeup" = 13, 
    "Screwball" = 14
    )

arsenal_p <- szn_pitch |>
  mutate("Primary" = 
           pmax(Fastball,Cutter,Split_Finger,Sinker,Slider,Curveball,Knuckle_Curve,Changeup,Screwball)) |>
  mutate("Secondary" = 
           apply(
             szn_pitch[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               second <- values_sorted[2]
               return(second)
             }
           )
  ) |> 
  mutate("Third" = 
           apply(
             szn_pitch[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               third <- values_sorted[3]
               return(third)
             }
           )
  ) |> 
  mutate("Fourth" = 
           apply(
             szn_pitch[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               fourth <- values_sorted[4]
               return(fourth)
             }
           )
  ) |> 
  mutate("Fifth" = 
           apply(
             szn_pitch[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               fifth <- values_sorted[5]
               return(fifth)
             }
           )
  )
```
  
```{r}
cor(arsenal_p$ERA_minus, arsenal_p[15:17])
```

### Pitch Grades

```{r}
szn_arsenal <- szn |>
  replace(is.na(szn),0) |>
  filter(Role == "SP", G > 15) |>
  select(4, 6, 9, 12, 97, 170:172, 174:177, 179, 180) |>
  rename(
    "Fastball" = 6, 
    "Cutter" = 7, 
    "Split_Finger" = 8, 
    "Sinker" = 9, 
    "Slider" = 10, 
    "Curveball" = 11, 
    "Knuckle_Curve" = 12, 
    "Changeup" = 13,
    "Screwball" = 14
    )

sa1 <- szn_arsenal |>
  mutate("Primary" = apply(szn_arsenal[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[1]])) |>
  mutate("Secondary" = apply(szn_arsenal[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[2]])) |>
  mutate("Third" = apply(szn_arsenal[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[3]])) |>
  mutate("Fourth" = apply(szn_arsenal[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[4]])) |>
  mutate("Fifth" = apply(szn_arsenal[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[5]])) |>
  select(1,2,3,5,15:19) |>
  pivot_longer(cols=c(5:9), names_to = "Arsenal", values_to = "Pitch")
```

```{r}
sa2 <- szn |>
  replace(is.na(szn),0) |>
  filter(Role == "SP", G > 15) |>
  select(4, 6, 9, 12, 97, 279, 282, 285, 288, 291, 294, 297, 300, 304) |>
  rename(
    "Changeup" = 6, 
    "Curveball" = 7, 
    "Fastball" = 8, 
    "Sinker" = 9, 
    "Slider" = 10, 
    "Knuckle_Curve" = 11, 
    "Cutter" = 12, 
    "Split_Finger" = 13, 
    "Forkball" = 14
    ) |> 
  select(1,2,3,5:14) |>
  pivot_longer(cols = c(5:13), names_to = "Pitch", values_to = "Stuff_plus")
```

```{r}
arsenal_grades <- left_join(sa1, sa2, by = c("NameASCII", "Season", "Role", "ERA_minus", "Pitch")) |>
  select(!6) |>
  pivot_wider(names_from = "Arsenal", values_from = "Stuff_plus") |>
  mutate(across(everything(), ~replace(., . == 0, NA)))
```

```{r}
cor(arsenal_grades$ERA_minus, arsenal_grades[5:7], use = "complete.obs")
```

## Single-Innings Relievers

### Pitch Percentages

```{r}
szn_pitch_rp <- szn |>
  replace(is.na(szn),0) |>
  mutate(Full_IP = trunc(IP),
         Partial_Outs = IP - trunc(IP),
         Outs = 3 * Full_IP + 10 * Partial_Outs,
         O_per_G = Outs/G) |>
  filter(Role == "RP", G > 15, O_per_G < 4) |>
  select(4, 6, 9, 12, 97, 170:172, 174:177, 179, 180) |>
  rename(
    "Fastball" = 6, 
    "Cutter" = 7, 
    "Split_Finger" = 8, 
    "Sinker" = 9, 
    "Slider" = 10, 
    "Curveball" = 11, 
    "Knuckle_Curve" = 12, 
    "Changeup" = 13, 
    "Screwball" = 14
    )

arsenal_p_rp <- szn_pitch_rp |>
  mutate("Primary" = 
           pmax(Fastball,Cutter,Split_Finger,Sinker,Slider,Curveball,Knuckle_Curve,Changeup,Screwball)) |>
  mutate("Secondary" = 
           apply(
             szn_pitch_rp[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               second <- values_sorted[2]
               return(second)
             }
           )
  ) |> 
  mutate("Third" = 
           apply(
             szn_pitch_rp[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               third <- values_sorted[3]
               return(third)
             }
           )
  ) |> 
  mutate("Fourth" = 
           apply(
             szn_pitch_rp[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               fourth <- values_sorted[4]
               return(fourth)
             }
           )
  ) |> 
  mutate("Fifth" = 
           apply(
             szn_pitch_rp[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               fifth <- values_sorted[5]
               return(fifth)
             }
           )
  )
```
  
```{r}
cor(arsenal_p_rp$ERA_minus, arsenal_p_rp[15:17])
```

### Pitch Grades

```{r}
szn_arsenalr <- szn |>
  replace(is.na(szn),0) |>
  mutate(Full_IP = trunc(IP),
         Partial_Outs = IP - trunc(IP),
         Outs = 3 * Full_IP + 10 * Partial_Outs,
         O_per_G = Outs/G) |>
  filter(Role == "RP", G > 15, O_per_G < 4) |>
  select(4, 6, 9, 12, 97, 170:172, 174:177, 179, 180) |>
  rename(
    "Fastball" = 6, 
    "Cutter" = 7, 
    "Split_Finger" = 8, 
    "Sinker" = 9, 
    "Slider" = 10, 
    "Curveball" = 11, 
    "Knuckle_Curve" = 12, 
    "Changeup" = 13,
    "Screwball" = 14
    )

ra1 <- szn_arsenalr |>
  mutate("Primary" = apply(szn_arsenalr[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[1]])) |>
  mutate("Secondary" = apply(szn_arsenalr[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[2]])) |>
  mutate("Third" = apply(szn_arsenalr[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[3]])) |>
  mutate("Fourth" = apply(szn_arsenalr[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[4]])) |>
  mutate("Fifth" = apply(szn_arsenalr[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[5]])) |>
  select(1,2,3,5,15:19) |>
  pivot_longer(cols=c(5:9), names_to = "Arsenal", values_to = "Pitch")
```

```{r}
ra2 <- szn |>
  replace(is.na(szn),0) |>
  mutate(Full_IP = trunc(IP),
         Partial_Outs = IP - trunc(IP),
         Outs = 3 * Full_IP + 10 * Partial_Outs,
         O_per_G = Outs/G) |>
  filter(Role == "RP", G > 15, O_per_G < 4) |>
  select(4, 6, 9, 12, 97, 279, 282, 285, 288, 291, 294, 297, 300, 304) |>
  rename(
    "Changeup" = 6, 
    "Curveball" = 7, 
    "Fastball" = 8, 
    "Sinker" = 9, 
    "Slider" = 10, 
    "Knuckle_Curve" = 11, 
    "Cutter" = 12, 
    "Split_Finger" = 13, 
    "Forkball" = 14
    ) |> 
  select(1,2,3,5:14) |>
  pivot_longer(cols = c(5:13), names_to = "Pitch", values_to = "Stuff_plus")
```

```{r}
arsenal_gradesr <- left_join(ra1, ra2, by = c("NameASCII", "Season", "Role", "ERA_minus", "Pitch")) |>
  select(!6) |>
  pivot_wider(names_from = "Arsenal", values_from = "Stuff_plus") |>
  mutate(across(everything(), ~replace(., . == 0, NA)))
```

```{r}
cor(arsenal_gradesr$ERA_minus, arsenal_gradesr[5:7], use = "complete.obs")
```

## Multi-Innings Relievers

### Pitch Percentages

```{r}
szn_pitch_lrp <- szn |>
  replace(is.na(szn),0) |>
  mutate(Full_IP = trunc(IP),
         Partial_Outs = IP - trunc(IP),
         Outs = 3 * Full_IP + 10 * Partial_Outs,
         O_per_G = Outs/G) |>
  filter(Role == "RP", G > 15, O_per_G > 4) |>
  select(4, 6, 9, 12, 97, 170:172, 174:177, 179, 180) |>
  rename(
    "Fastball" = 6, 
    "Cutter" = 7, 
    "Split_Finger" = 8, 
    "Sinker" = 9, 
    "Slider" = 10, 
    "Curveball" = 11, 
    "Knuckle_Curve" = 12, 
    "Changeup" = 13, 
    "Screwball" = 14
    )

arsenal_p_lrp <- szn_pitch_lrp |>
  mutate("Primary" = 
           pmax(Fastball,Cutter,Split_Finger,Sinker,Slider,Curveball,Knuckle_Curve,Changeup,Screwball)) |>
  mutate("Secondary" = 
           apply(
             szn_pitch_lrp[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               second <- values_sorted[2]
               return(second)
             }
           )
  ) |> 
  mutate("Third" = 
           apply(
             szn_pitch_lrp[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               third <- values_sorted[3]
               return(third)
             }
           )
  ) |> 
  mutate("Fourth" = 
           apply(
             szn_pitch_lrp[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               fourth <- values_sorted[4]
               return(fourth)
             }
           )
  ) |> 
  mutate("Fifth" = 
           apply(
             szn_pitch_lrp[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) {
               values_sorted <- sort(x, decreasing=T)
               fifth <- values_sorted[5]
               return(fifth)
             }
           )
  )
```
  
```{r}
cor(arsenal_p_lrp$ERA_minus, arsenal_p_lrp[15:17])
```

### Pitch Grades

```{r}
szn_arsenallr <- szn |>
  replace(is.na(szn),0) |>
  mutate(Full_IP = trunc(IP),
         Partial_Outs = IP - trunc(IP),
         Outs = 3 * Full_IP + 10 * Partial_Outs,
         O_per_G = Outs/G) |>
  filter(Role == "RP", G > 15, O_per_G > 4) |>
  select(4, 6, 9, 12, 97, 170:172, 174:177, 179, 180) |>
  rename(
    "Fastball" = 6, 
    "Cutter" = 7, 
    "Split_Finger" = 8, 
    "Sinker" = 9, 
    "Slider" = 10, 
    "Curveball" = 11, 
    "Knuckle_Curve" = 12, 
    "Changeup" = 13,
    "Screwball" = 14
    )

lra1 <- szn_arsenallr |>
  mutate("Primary" = apply(szn_arsenallr[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[1]])) |>
  mutate("Secondary" = apply(szn_arsenallr[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[2]])) |>
  mutate("Third" = apply(szn_arsenallr[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[3]])) |>
  mutate("Fourth" = apply(szn_arsenallr[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[4]])) |>
  mutate("Fifth" = apply(szn_arsenallr[,c("Fastball","Cutter","Split_Finger","Sinker","Slider","Curveball","Knuckle_Curve","Changeup","Screwball")], 1, function(x) names(x)[order(x, decreasing = T)[5]])) |>
  select(1,2,3,5,15:19) |>
  pivot_longer(cols=c(5:9), names_to = "Arsenal", values_to = "Pitch")
```

```{r}
lra2 <- szn |>
  replace(is.na(szn),0) |>
  mutate(Full_IP = trunc(IP),
         Partial_Outs = IP - trunc(IP),
         Outs = 3 * Full_IP + 10 * Partial_Outs,
         O_per_G = Outs/G) |>
  filter(Role == "RP", G > 15, O_per_G > 4) |>
  select(4, 6, 9, 12, 97, 279, 282, 285, 288, 291, 294, 297, 300, 304) |>
  rename(
    "Changeup" = 6, 
    "Curveball" = 7, 
    "Fastball" = 8, 
    "Sinker" = 9, 
    "Slider" = 10, 
    "Knuckle_Curve" = 11, 
    "Cutter" = 12, 
    "Split_Finger" = 13, 
    "Forkball" = 14
    ) |> 
  select(1,2,3,5:14) |>
  pivot_longer(cols = c(5:13), names_to = "Pitch", values_to = "Stuff_plus")
```

```{r}
arsenal_gradeslr <- left_join(lra1, lra2, by = c("NameASCII", "Season", "Role", "ERA_minus", "Pitch")) |>
  select(!6) |>
  pivot_wider(names_from = "Arsenal", values_from = "Stuff_plus") |>
  mutate(across(everything(), ~replace(., . == 0, NA)))
```

```{r}
cor(arsenal_gradeslr$ERA_minus, arsenal_gradeslr[5:7], use = "complete.obs")
```

# Pitch Arsenal Linear Regression

## Starter

```{r}
sp_mod2 <- lm(ERA_minus ~ Primary + Secondary + Third, data=arsenal_grades)
summary(sp_mod2)
```

## SI Reliever

```{r}
rp_mod1 <- lm(ERA_minus ~ Primary + Secondary + Third, data=arsenal_gradesr)
summary(rp_mod1)

rp_mod2 <- lm(ERA_minus ~ Primary + Secondary, data=arsenal_gradesr)
summary(rp_mod2)
```

## L Reliever

```{r}
lrp_mod1 <- lm(ERA_minus ~ Primary + Secondary + Third, data=arsenal_gradeslr)
summary(lrp_mod1)

lrp_mod2 <- lm(ERA_minus ~ Primary + Secondary, data=arsenal_gradeslr)
summary(lrp_mod2)

lrp_mod3 <- lm(ERA_minus ~ Primary, data=arsenal_gradeslr)
summary(lrp_mod3)
```

# Plus Pitches

```{r}
plus_pitches <- szn |>
  replace(is.na(szn),0) |>
  filter(G > 15) |>
  select(4, 6, 9, 12, 97, 279, 282, 285, 288, 291, 294, 297, 300, 304) |>
  rename(
    "Changeup" = 6, 
    "Curveball" = 7, 
    "Fastball" = 8, 
    "Sinker" = 9, 
    "Slider" = 10, 
    "Knuckle_Curve" = 11, 
    "Cutter" = 12, 
    "Split_Finger" = 13, 
    "Forkball" = 14
    ) |>
  mutate(
    plus_pitches = rowSums(across(6:14, ~ . > 110, .names = "V{.col}")),
    ERA_dummy = ERA_minus < 100
    )

ggplot(data=plus_pitches,
    aes(x=plus_pitches, y=ERA_minus)
  ) + geom_smooth(method="lm") + facet_wrap(~Role) +
  labs(title = "ERA-minus vs Number of +Pitches",
       x="Number of Plus Pitches",
       y="ERA_minus")
```

```{r}
plus_pitches |>
  group_by(plus_pitches, Role, ERA_dummy) |>
  summarize(count = n()) |>
  ggplot(
       aes(x=plus_pitches, y=count, fill=ERA_dummy)) +
    geom_bar(stat="identity") + facet_wrap(~Role) +
  labs(title = "Pitchers with Above-Average ERA-minus",
       x= "Number of Plus Pitches",
       y= "Number of Pitchers",
       fill="ERA- < 100")
```

# wOBA Across an Outing

```{r}
duration <- pitches |>
  select(player_name, woba_value, role_key, pitch_number_appearance, pitcher_at_bat_number, times_faced) |>
  filter(!is.na(woba_value)) |>
  group_by(role_key, pitch_number_appearance) |>
  summarize(mean_woba = mean(woba_value), count = n())
```

```{r}
duration |>
  ggplot(
    aes(x=pitch_number_appearance, y=mean_woba, color=role_key)
  ) + geom_smooth(method = "loess", se=F) +
  labs(tite = "Mean wOBA vs Number of Pitches",
       x="Pitch Count",
       y="mean wOBA",
       color="Role")
```

# Maintaining Velocity

```{r}
results <- pitches |>
  filter(pitch_type == "FF") |>
  group_by(pitch_type, game_date, player_name, role_key, pitcher_at_bat_number) |>
  summarize(mean_velo = mean(release_speed)) |>
  pivot_wider(names_from = pitcher_at_bat_number, names_sort = T, values_from = mean_velo) |>
  mutate(across(starts_with("2") : ends_with("35"), ~ . - `1`, .names = "{.col}")) |>
  group_by(player_name, role_key) |>
  summarize(across(c(4:37), mean, na.rm = TRUE)) |>
  pivot_longer(cols = c(3:36), names_to = "Pitch", values_to = "Difference") |>
  mutate(Pitch = as.numeric(Pitch)) |> 
  na.omit()
```

```{r}
ggplot(data=results,
       aes(x=Pitch, y=Difference)
  ) + geom_point() + geom_smooth(method = "lm", se=F) + facet_wrap(~role_key) +
  labs(title = "Four-Seam Fastball Velocity Average Difference", x="At-Bat", y="Difference in Velo from First Pitch")
```

```{r}
results_cu <- pitches |>
  filter(pitch_type == "FC") |>
  group_by(pitch_type, game_date, player_name, role_key, pitcher_at_bat_number) |>
  summarize(mean_velo = mean(release_speed)) |>
  pivot_wider(names_from = pitcher_at_bat_number, names_sort = T, values_from = mean_velo) |>
  mutate(across(starts_with("2") : ends_with("35"), ~ . - `1`, .names = "{.col}")) |>
  group_by(player_name, role_key) |>
  summarize(across(c(4:37), mean, na.rm = TRUE)) |>
  pivot_longer(cols = c(3:36), names_to = "Pitch", values_to = "Difference") |>
  mutate(Pitch = as.numeric(Pitch)) |> 
  na.omit()
```

```{r}
ggplot(data=results_cu,
       aes(x=Pitch, y=Difference)
  ) + geom_point() + geom_smooth(method = "lm", se=F) + facet_wrap(~role_key) +
  labs(title = "Cutter Velocity Average Difference", x="At-Bat", y="Difference in Velo from First Pitch")
```

```{r}
results_si <- pitches |>
  filter(pitch_type == "SI") |>
  group_by(pitch_type, game_date, player_name, role_key, pitcher_at_bat_number) |>
  summarize(mean_velo = mean(release_speed)) |>
  pivot_wider(names_from = pitcher_at_bat_number, names_sort = T, values_from = mean_velo) |>
  mutate(across(starts_with("2") : ends_with("35"), ~ . - `1`, .names = "{.col}")) |>
  group_by(player_name, role_key) |>
  summarize(across(c(4:37), mean, na.rm = TRUE)) |>
  pivot_longer(cols = c(3:36), names_to = "Pitch", values_to = "Difference") |>
  mutate(Pitch = as.numeric(Pitch)) |> 
  na.omit()
```

```{r}
ggplot(data=results_si,
       aes(x=Pitch, y=Difference)
  ) + geom_point() + geom_smooth(method = "lm", se=F) + facet_wrap(~role_key) +
  labs(title = "Sinker Velocity Average Difference", x="At-Bat", y="Difference in Velo from First Pitch")
```

# Candidates for New Roles

## Rafael Montero 

```{r}
szn |>
  replace(is.na(szn),0) |>
  filter(NameASCII == "Rafael Montero") |>
  select(4, 6, 9, 12, 97, 170, 174, 175, 179, 279, 285, 288, 291)
```

```{r}
results |>
  filter(player_name == "Montero, Rafael") |>
  arrange(-Pitch)
```

## Stephen Strasburg

```{r}
szn |>
  replace(is.na(szn),0) |>
  filter(NameASCII == "Stephen Strasburg") |>
  select(4, 6, 9, 12, 97, 170, 174, 176, 179, 279, 282, 285, 288)
```

```{r}
results |>
  filter(player_name == "Strasburg, Stephen") |>
  arrange(-Pitch)
```

```{r}
results_si |>
  filter(player_name == "Strasburg, Stephen") |>
  arrange(-Pitch)
```